### 2023-11-23 main.tex
^-^

	\Omega=\bigotimes_{x=1}^{L} \Omega_{x}

@@

Here $r=(r_{1},\ldots,r_{N})$ and  $\epsilon=(\epsilon_{1},\ldots,\epsilon_{N})$ such that $\forall a\in\{1,\ldots,N\}$  $r_{a}\in \{1,\ldots,\nu\}$ and $\epsilon_{a}\in \{0,\ldots,2\nu\}$, under the constraints 

@@

Therefore 

@@

In the thermalized bulk generator particles on a bond $x$ and $x+1$ are redistribute according to the measure $\nu_{TH}(r|\epsilon)$, while in the thermalized boundary generators the configuration at site $1$ (and $L$) is replaced by $r$, according to a rate depending on $r$ itself and on the reservoirs parameters. 
\subsubsection{The dual thermalized process}
The thermalized dual process $(\bm{\xi}(t))_{t\geq 0}$ is defined on the enlarged configuration space $\widetilde{\Omega}$ introduced in \eqref{dualStateSpace}. Its generator reads

@@

where $\mathcal{L}^{TH}_{x,x+1}$ is again \eqref{bulkThermalized} and $\widetilde{\mathcal{L}}_{\text{left}}^{TH}$ and $\widetilde{\mathcal{L}}_{\text{right}}^{TH}$ are absorbing with rate $1$:

@@

This means that they remove all dual particles at site $1$ (and $L$) and put these particles in the extra-site $0$ (and $L+1$) with rate 1. The duality function is the same as the original process, i.e.

@@

The proof duality result for the thermalized multi-species stirring process is standard. For the sake of completeness we report it in Appendix \ref{appendix-dualityThermalized}.

@@

@@

\begin{remark}: observe that, up to a specific choice of the $\alpha$'s, the left boundary generator $\mathcal{L}_{\text{left}}^{rd}$ has the form of the one of the multi-species stirring process \eqref{siteGenerator}. 

@@

On the occupation variable $n_{a}^{L}$ the action is similar. \\ In the spirit of \cite{casini2022uphill} we define the average density occupation of species $a$ at site $x$ as
where $\mathbb{E}_{\mu_{0}}$ denotes the expectation with respect to the law of the Markov process $(\bm{n}(t))_{t\geq 0}$ initialized with arbitrary initial measure $\mu_{0}$. These average occupations evolve according to the master equation
Therefore, using the action of $\mathcal{L}^{rd}$ expressed by \eqref{DifferenceEquation} and \eqref{actionBDLine} we obtain the desired system of $N-1$ difference-differential equations \eqref{NSpeciesRD}. For these reasons, this model is a generalization to $N$ species and to maximal occupation $\nu$ of the one studied in \cite{casini2022uphill}. This answers the first question we addressed to at the beginning of this subsection. 

@@

	p_{a}p_{b}=p_{h_{c}(a)}p_{h_{c}(b)}
By the arbitrariness of $a,b,$ it is true if and only if
\begin{equation}\label{p-equal}
	p_{a}=p\qquad \forall a\in \{1,\ldots,N-1\}
while $p_{N}=1-\sum_{a=1}^{N-1}p_{a}$\\
For the boundary generators $\mathcal{L}_{\text{left}}^{rd}$ and $\mathcal{L}_{\text{right}}^{rd}$ we have that, analogously to what happens for the multispecies stirring process, the reservoirs parameters must be such that $\beta_{a}=\alpha_{a}$ and, due to \eqref{p-equal}, we also need $\alpha_{a}=\alpha$ for all $a\in\{1,\ldots,N-1\}$. 
\subsubsection{Duality for the reaction-diffusion process}
In this subsection we formulate duality for the multi-species stirring process $(\bm{n}(t))_{t\geq 0}$, defined by the generator \eqref{RDGenerator}.
The dual  process $(\bm{\xi}(t))_{t\geq 0}$ takes values in the enlarged state space $\widetilde{\Omega}$ defined in \eqref{dualStateSpace}
%	\widetilde{\Omega}=\mathbb{N}_{0}^{N}\times  \bigtimes_{x\in V} \Omega_{x}\times \mathbb{N}_{0}^{N}
%We associate an ``extra site'', denoted by $0$ and $L+1$ attached to $1$ and $L$ respectively,
%where dual particles will accumulate in the course of time. 
%and we denote its configurations by $\bm{\xi} \in \widetilde\Omega$.
%	\bm{\xi}=\left(\xi_{1}^{0},\ldots,\xi_{N}^{0}\right)\otimes\left(\xi_{1}^{x},\ldots,\xi_{N}^{x}\right)^{\otimes L}\otimes \left(\xi_{1}^{L+1},\ldots,\xi_{N}^{L+1}\right)
%where the component $\xi_{a}^{x}$ is interpreted as the number of dual particles of type $a\in \{1,\ldots,N\}$ at site $x$, 
%and the component $\xi_{a}^{0},\xi_{a}^{L+1}$  give the number at the extra sites. \\
	D(\bm{n},\bm{\xi})=\prod_{a=1}^{N-1}\left(\frac{\alpha_{a}}{|\alpha|}\right)^{\xi_{a}^{0}}\left(\prod_{x=1}^{L}\frac{(\nu -\sum_{a=1}^{N-1}\xi_{a}^{x})!}{\nu!}\prod_{a=1}^{N-1}\frac{n_{a}^{x}!}{(n_{a}^{x}-\xi_{a}^{x})!}\right)\prod_{a=1}^{N-1}\left(\frac{\beta_{a}}{|\beta|}\right)^{\xi_{a}^{L+1}}\dt
%where we denote the \textit{density} of the species $a\in \{2,\ldots,N\}$ imposed by the reservoir at site $x\in V$  by 
%	\rho_{a}^{x}:=\frac{\alpha_{a}^{x}}{|\alpha^{x}|}
The dual generator now reads
\begin{equation}\label{DualGeneratorRD}
	\widetilde{\mathcal{L}}_{\text{left}}^{rd}=\widetilde{\mathcal{L}}^{rd}+\sum_{x=1}^{L-1}\mathcal{L}_{x,x+1}^{rd}+\widetilde{\mathcal{L}}_{\text{right}}^{rd}
$\mathcal{L}_{x,x+1}^{rd}$ is defined in \eqref{edgeGeneratorRD} and, $\widetilde{\mathcal{L}}_{\text{left}}^{rd}$ and $\widetilde{\mathcal{L}}_{\text{right}}^{rd}$ are again the absorbing dual generators $\widetilde{\mathcal{L}}_{x}$ defined in \eqref{siteDualGenerator}, with the specific choice of the reservoirs parameters \eqref{boundaryParamRD}. This answer to the second question addressed at the beginning of this Section. 
\paragraph{Proof of duality:}
we only prove duality for the bulk generator, since the boundary generators are the same as in Section \ref{Subsection-ss-nonI}. 
In the spirit of Section \ref{sectionDuality}, it is convenient to write the Hamiltonian of the process, that is defined through the basis elements \eqref{actionE} of the $gl(N)$ Lie algebra as 

@@

Here, $\mathcal{H}_{x,x+1}$ and $H_{\text{left}},H_{\text{right}}$ are \eqref{edgeHamiltonian}  and \eqref{siteHamiltonian} (with the choice of the boundary parameters done in \eqref{boundaryParamRD}) respectively, while

@@

%	(\mathcal{H}_{x,x+1}^{rd})^{T}D&=D\mathcal{H}_{x,x+1}^{rd}\label{edgeDualityRelationRD}
%but it is enough of showing 

>>>>

	\Omega_{L}=\bigotimes_{x=1}^{L} \Omega_{x}

@@

Here $r=(r_{1},\ldots,r_{N})$ and  $\epsilon=(\epsilon_{1},\ldots,\epsilon_{N})$ such that $\forall a\in\{1,\ldots,N\}$  $r_{a}\in \{0,\ldots,\nu\}$ and $\epsilon_{a}\in \{0,\ldots,2\nu\}$, under the constraints 

@@

Therefore the thermalization measure reads

@@

In the thermalized bulk generator particles on a bond $x$ and $x+1$ are redistribute according to the measure $\nu_{TH}(r|\epsilon)$, while in the thermalized boundary generators the configuration at site $1$ ($L$) is replaced by $r$, according to a rate depending on $r$ itself and on the reservoirs parameters. 
\subsubsection{The dual thermalized process}\label{subsection-dualTHM}
Here we state the duality result for the thermalized multi-species stirring process with boundary driving. In the following, the dual configuration space is the one defined in \eqref{dualStateSpace} and adapted to the chain of size $L$ and with two extra-site $\{0,L+1\}$ attached to $\{1,L\}$ respectively. To recall it we denote this dual space by $\widetilde{\Omega}_{L}$.
\begin{proposition}[Duality for the thermalized multi-species stirring process]\label{proposition-dualityTHM}
The thermalized multi-species stirring process $(\bm{n}(t))_{t\geq 0}$ on the state space $\Omega_{L}$, with generator $\mathcal{L}^{TH}$ defined in \eqref{thermalizedGenerator} is dual to the process $(\bm{\xi}(t))_{t\geq 0}$ on the enlarged configuration space $\widetilde{\Omega}_{L}$, with dual generator

@@

where $\mathcal{L}^{TH}_{x,x+1}$ is \eqref{bulkThermalized} and $\widetilde{\mathcal{L}}_{\text{left}}^{TH}$ and $\widetilde{\mathcal{L}}_{\text{right}}^{TH}$ are absorbing with rate $1$:

@@

The duality function is the same of the multi-species stirring process, i.e. 

@@

(cf. \eqref{dualityElements}).
\end{proposition}
This dual process is absorbing since the dual boundary generator $\mathcal{L}_{\text{left}}^{TH}$ ($\mathcal{L}_{\text{left}}^{TH}$) remove all dual particles at site $1$ ($L$) and put these particles in the extra-site $0$ ($L+1$) with rate 1.
The proof of proposition \ref{proposition-dualityTHM} is standard. For the sake of completeness we report it in Appendix \ref{appendix-dualityThermalized}.

@@

We define the average density occupation of species $a$ at site $x$ as
where $\mathbb{E}_{\mu_{0}}$ denotes the expectation with respect to the law of the Markov process $(\bm{n}(t))_{t\geq 0}$ initialized with arbitrary initial measure $\mu_{0}$. These average occupations evolve according to the evolution equation
\begin{equation}\label{EE-density}

@@

\begin{remark}: observe that, up to a specific choice of the $\alpha$'s ($\beta$'s), the left (right) boundary generator $\mathcal{L}_{\text{left}}^{rd}$ ($\mathcal{L}_{\text{right}}^{rd}$) has the form of the one of the multi-species stirring process \eqref{siteGenerator}. 

@@

On the occupation variable $n_{a}^{L}$ the action is similar. Therefore, using the evolution equation \eqref{EE-density} for the average densities and  the action of $\mathcal{L}^{rd}$ expressed by \eqref{DifferenceEquation} and \eqref{actionBDLine} we obtain the desired system of $N-1$ difference-differential equations \eqref{NSpeciesRD}. For these reasons, this model is a generalization to $N$ species and to maximal occupation $\nu$ of the one studied in \cite{casini2022uphill}. This answers the first question we addressed to at the beginning of this section. 

@@

\begin{equation}
	p_{a}p_{b}=p_{h_{c}(a)}p_{h_{c}(b)}
\end{equation}
By the arbitrariness of $a,b,$ it is true if and only if
\begin{equation}\label{p-equal}
	p_{a}=p\qquad \forall a\in \{1,\ldots,N-1\}
\end{equation}
while $p_{N}=1-\sum_{a=1}^{N-1}p_{a}$\\
For the boundary generators $\mathcal{L}_{\text{left}}^{rd}$ and $\mathcal{L}_{\text{right}}^{rd}$ we have that, analogously to what happens for the multispecies stirring process, the reservoirs parameters must be such that $\beta_{a}=\alpha_{a}$ and, due to \eqref{p-equal}, we also need $\alpha_{a}=\alpha$ for all $a\in\{1,\ldots,N-1\}$. 
\end{comment}
\subsubsection{Duality for the reaction-diffusion process}
In this subsection we formulate duality for the multi-species reaction diffusion stirring process. This answer to the second question addressed at the beginning of this Section.  The dual process is defined on the enlarged state space $\widetilde{\Omega}_{L}$, introduced in Section \ref{subsection-dualTHM}.  
\begin{proposition}[Duality for the reaction-diffusion process]\label{propositin-duality-RD}
The reaction-diffusion multi-species stirring process $(\bm{n}(t))_{t\geq 0}$, on the state space $\Omega_{L}$, with generator $\mathcal{L}^{rd}$ defined in \eqref{RDGenerator} is dual to the process $(\bm{\xi}(t))_{t\geq 0}$ on the state space $\widetilde{\Omega}_{L}$ with dual generator
\begin{equation}\label{DualGeneratorRD}
	\widetilde{\mathcal{L}}_{\text{left}}^{rd}=\widetilde{\mathcal{L}}^{rd}+\sum_{x=1}^{L-1}\mathcal{L}_{x,x+1}^{rd}+\widetilde{\mathcal{L}}_{\text{right}}^{rd}
\end{equation}
where 
$\mathcal{L}_{x,x+1}^{rd}$ is defined in \eqref{edgeGeneratorRD} and, $\widetilde{\mathcal{L}}_{\text{left}}^{rd}$ and $\widetilde{\mathcal{L}}_{\text{right}}^{rd}$ are the absorbing dual generators $\widetilde{\mathcal{L}}_{x}$ defined in \eqref{siteDualGenerator} and acting at sites $1$ and $L$, with the specific choice of the reservoirs parameters \eqref{boundaryParamRD}. The duality function
is the same of the stirring process, i.e.
\begin{equation}
	D(\bm{n},\bm{\xi})=\prod_{a=1}^{N-1}\left(\frac{\alpha_{a}}{|\alpha|}\right)^{\xi_{a}^{0}}\left(\prod_{x=1}^{L}\frac{(\nu -\sum_{a=1}^{N-1}\xi_{a}^{x})!}{\nu!}\prod_{a=1}^{N-1}\frac{n_{a}^{x}!}{(n_{a}^{x}-\xi_{a}^{x})!}\right)\prod_{a=1}^{N-1}\left(\frac{\beta_{a}}{|\beta|}\right)^{\xi_{a}^{L+1}}
\end{equation}
(cf. \eqref{dualityElements}).
\end{proposition} 
%\begin{equation}
%	\widetilde{\Omega}=\mathbb{N}_{0}^{N}\times  \bigtimes_{x\in V} \Omega_{x}\times \mathbb{N}_{0}^{N}
%\end{equation}
%We associate an ``extra site'', denoted by $0$ and $L+1$ attached to $1$ and $L$ respectively,
%where dual particles will accumulate in the course of time. 
%and we denote its configurations by $\bm{\xi} \in \widetilde\Omega$.
%\begin{equation}
%	\bm{\xi}=\left(\xi_{1}^{0},\ldots,\xi_{N}^{0}\right)\otimes\left(\xi_{1}^{x},\ldots,\xi_{N}^{x}\right)^{\otimes L}\otimes \left(\xi_{1}^{L+1},\ldots,\xi_{N}^{L+1}\right)
%\end{equation}
%where the component $\xi_{a}^{x}$ is interpreted as the number of dual particles of type $a\in \{1,\ldots,N\}$ at site $x$, 
%and the component $\xi_{a}^{0},\xi_{a}^{L+1}$  give the number at the extra sites. \\
%where we denote the \textit{density} of the species $a\in \{2,\ldots,N\}$ imposed by the reservoir at site $x\in V$  by 
%\begin{equation}
%	\rho_{a}^{x}:=\frac{\alpha_{a}^{x}}{|\alpha^{x}|}
%\end{equation}

\paragraph{Proof of Proposition \ref{propositin-duality-RD}:}
we only prove duality for the bulk generator, since the boundary generators are the same of the multi-species stirring process. In the spirit of Section \ref{sectionDuality}, it is convenient to write the Hamiltonian of the process, that is defined through the basis elements \eqref{actionE} of the $gl(N)$ Lie algebra as 

@@

Here, $\mathcal{H}_{x,x+1}$ is  \eqref{edgeHamiltonian}  and $H_{\text{left}},H_{\text{right}}$ is \eqref{siteHamiltonian} (with the choice of the boundary parameters done in \eqref{boundaryParamRD}). We, then have that 

@@
